---
title: "Assignment 6"
output:
  html_document:
    df_print: paged
---


```{r}
## Load up some packages we will be using for Decision Trees
library(caTools)
library(rpart)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(caret) 
```


```{r}
diabetes <- read.csv("diabetes.csv")
```

1. Split your data using a 2/3 training and 1/3 test split
```{r}
set.seed(5)
d_sample <- sample.split(diabetes$Outcome,SplitRatio=0.67)
diabetes_train <- subset(diabetes,d_sample==TRUE)
diabetes_test <- subset(diabetes,d_sample==FALSE)
```


2. Create a full decision tree with your Training Data. Outcome is your class label and all other variables can be used as attributes of your model
```{r}
diabetes_tree <- rpart(Outcome ~ ., data=diabetes_train, method = "class",
                     parms = list(split = 'information'),cp=-1)
diabetes_tree
```


3. Create the graphic of the decision tree you just made
```{r}
fancyRpartPlot(diabetes_tree)
```



4. Test the Accuracy of this model of with Test Data
```{r}
pred <-predict(diabetes_tree,diabetes_test,type = "class")
conf <- table(pred,diabetes_test$Outcome)
confusionMatrix(conf)
confusionMatrix(conf)$byClass
```


5. Create a post-pruned decision tree with your Training Data with the graphic. Use 
```{r}
printcp(diabetes_tree)
plotcp(diabetes_tree)
cp = diabetes_tree$cptable[which.min(diabetes_tree$cptable[,"xerror"]),"CP"]
```

## Fit the Model
```{r}
post_diabetes_tree <- rpart(Outcome ~ ., data=diabetes_train, method = "class",
             parms = list(split = 'information'),cp=cp)
post_diabetes_tree
```

```{r}
fancyRpartPlot(post_diabetes_tree)
```

6. Test the accuracy of your post-pruned tree with the Test Data. Did we improve the fit of the model by pruning?

We improve the models by 5.5 percent. The first model had about 70% accuracy, and the post pruning model about 75.5% accuracy.

## Make the Predictions
```{r}
post_pred<-predict(post_diabetes_tree,diabetes_test,type = "class")
```

## Measures of Fit
```{r}
post_conf <- table(post_pred,diabetes_test$Outcome)
confusionMatrix(post_conf)
confusionMatrix(post_conf)$byClass
```
