---
title: "R Notebook"
output: html_notebook
---

```{r setup}
install.packages("caTools")
install.packages("rpart")
install.packages("rattle")
install.packages("rpart.plot")
install.packages("RColorBrewer")
install.packages("caret")
```



```{r}
## Load up some packages we will be using for Decision Trees
library(caTools)
library(rpart)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(caret) 
```



## This tutorial is built with the diabetes.csv dataset. Load that before starting
```{r}
loan <- read.csv("../case_study_1/loan.csv")
```

## Split the data into a training set and test test
```{r}
set.seed(5)
sample <- sample.split(loan$Default,SplitRatio=0.70)
train <- subset(loan,sample==TRUE)
test <- subset(loan,sample==FALSE)
```

## Creating the Full Tree Model
```{r}
tree <- rpart(Default ~ ., data=train, method = "class",
                     parms = list(split = 'information'),cp=-1)
tree
fancyRpartPlot(tree)
```



## Calculate the Predictions from the test data
```{r}
pred <-predict(tree,test,type = "class")
```


## Calculate the Confusion Matrix and fit measures
```{r}
conf <- table(pred,test$Default)
confusionMatrix(conf)
confusionMatrix(conf)$byClass
```











## Post-Pruning Method
## Find the CP
```{r}
printcp(tree)
plotcp(tree)
cp = tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
```



## Fit the Model
```{r}
post_tree <- rpart(Default ~ ., data=train, method = "class",
             parms = list(split = 'information'),cp=cp)
post_tree
fancyRpartPlot(post_tree)
```



## Make the Predictions
```{r}
post_pred<-predict(post_tree,test,type = "class")
```



## Measures of Fit
```{r}
post_conf <- table(post_pred,test$Default)
confusionMatrix(post_conf)
confusionMatrix(post_conf)$byClass
```



## How does it stack up to the Logistic Model?
```{r}
fit <- glm(Default~Checking_amount+Term+Credit_score+
             Education_loan+Saving_amount+Age,data = test, family = binomial)
fit_pred <- round(predict(fit,test,type = "response"))
fit_conf <- table(fit_pred,test$Default)
confusionMatrix(fit_conf)
confusionMatrix(fit_conf)$byClass
```

```{r}
confusionMatrix(conf)
confusionMatrix(pre_conf)
confusionMatrix(post_conf)
confusionMatrix(fit_conf)
```



